Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression

### 2018-08-15 revision_1/paper.tex
revisions to the proposal density derivation sections

For the MCMC algorithm,  conditioning statistics are assumed to be of the form $T(\by) = (\bb(X, \by), s(X, \by)$ where $\bb(X, \by)\in \mathbb R^{p}$ is an estimator for the regression coefficients and $s(X, \by)\in \{0\} \cup {\mathbb R}^+$ is an estimator of the scale. Throughout, observed data and summary statistic is denoted by $\by_{obs}$ and $T(\by_{obs})=(\bb(X, \by_{obs}), s(X, \by_{obs}))$, respectively.

% In particular, we demonstrate the method using simultaneous M-estimators defined by \eqref{Mest} for the one-sample setting and easily extended to the linear model in \eqref{LinearModel}. The estimator of the \green{regression} coefficients is denoted by $\bb(X, \by)\in \mathbb R^{p}$ and that of the scale by \green{\sout{$s(X, \by)\in \mathbb R$} $s(X, \by)\in \{0\} \cup {\mathbb R}^+$} . Thus, $T(\by)=(\bb(X, \by), s(X, \by))$.

Several conditions are imposed on the model and statistic to ensure validity of the MCMC algorithm:
\begin{itemize}
\labitem{C1}{fullRank} The $n \times p$ design matrix, $X$, whose $i^{th}$ row is $x_i^\top$, 
is of full column rank.  
\labitem{C2}{supReal} The $\epsilon_i$ are a random sample from some distribution which has a density with 
respect to Lebesgue measure on the real line and for which the support is the real line.  
\labitem{C3}{asb}$\bb(X,\by)$ is almost surely continuous and differentiable with respect to $\by$.  
\labitem{C4}{as} $s(X,\by)$ is almost surely positive, continuous, and differentiable with respect to $\by$.  
\labitem{C5}{regEq} $\bb(X,\by+X\bv)=\bb(X,\by)+\bv \ \ \text{for  all}\ \bv\in\mathbb{R}^p$. 
\labitem{C6}{scaleEqReg} $\bb(X,a\by)=a\bb(X,\by)\ \ \ \text{for all constants } a$.  
\labitem{C7}{regIn} $s(X,\by+X\bv)=s(X,\by) \ \ \text{for all}\ \bv\in\mathbb{R}^p$.  
\labitem{C8}{scaleEq2Reg} $s(X, a\by)=|a|s(X,\by) \ \ \text{for all constants } a$.  
\end{itemize}
Properties \ref{regEq} and \ref{scaleEqReg} of $\bb$ are called
\textit{regression} and \textit{scale equivariance},
respectively.  Properties \ref{regIn} and \ref{scaleEq2Reg} of $s$ are called \textit{regression invariance}
and \textit{scale equivariance}. 
Many estimators satisfy the above properties, including simultaneous M-estimators \citep{huber2009, maronna2006} for which the R package \texttt{brlm} (\texttt{github.com/jrlewi/brlm}) is available to implement the MCMC. Further software development is required to extend the MCMC implementation beyond these M-estimators. The package also implements the direct computational methods described in \cite{lewis2014}. These methods are effective in lower dimensional problems and were used in several of the examples in Section \ref{illustrations}.

@@

%The general style of algorithm that we present relies on the decomposition of the sampling density in (\ref{FullLikelihood})
%into one piece involving only $T(\by)$ and a second piece for the
%complete data $\by$ given $T(\by)$. 
%Relying on the modularity of MCMC algorithms, we construct a data augmented Gibbs sampler targeting $\pi(\bth, \by | T(\by)=T(\by_{obs}))$, the joint posterior of $\bth$ and $\by$ conditioned on the summary statistic $T(\by_{obs})$ for observed data vector $\by_{obs}$. The Gibbs sampler iteratively samples from the full conditionals written as $\pi(\bth|\by T(\by)=T(\by_{obs})$ and $f(\by|\bth, T(\by)=T(\by_{obs})$
% \textcolor{blue}{some major changes from here to the end of the section}
The general style of algorithm that we present is a data augmented
MCMC algorithm targeting $f(\bth, \by |
T(\by)=T(\by_{obs}))$, the joint distribution of $\bth$ and the full
data given the summary statistic $T(\by_{obs})$.  % with $\by_{obs}$  representing the observed complete data. 
 The Gibbs sampler \citep{gelfand1990} iteratively samples from the
 full conditionals $\pi(\bth|\by, T(\by)=T(\by_{obs}))$ and
 $f(\by|\bth, T(\by)=T(\by_{obs}))$.  
\green{Alternative algorithms substitute other generations for the full conditionals, such as the Metropolis-Hastings propose and accept/reject framework.  Our algorithms are based on the decomposition of the MCMC algorithm into conventional full data sampling steps where $\bth$ is generated given $\by$, and steps that allow us to complete the data set where $\by$ is generated given $\bth$ and $T(\by)$.  The latter generation relies on the conditioning set $\mathcal{A}=\{\by\in \mathbb{R}^n|T(\by)=T(\by_{obs})\}$.  The set $\mathcal{A}$ may contain more than one element (i.e., it is not degenerate) when $T(\by) : \by \rightarrow T(\by)$ is not a $1-1$ function.}
%\red{\sout{, where the two
%     conditioning sets are not degenerate (that is, $\by$ is such that
%     $T(\by)=T(\by_{obs})$).}}  \red{Unless $T(\by)=\by$, the set
%   $\mathcal{A}=\{\by\in \mathbb{R}^n|T(\by)=T(\by_{obs})\}$
%   is not degenerate, containing more elements than $\by_{obs}$ only.}

When $\by$ has the summary statistic \sout{$T(\by)$ matching $T(\by_{obs})$,} \green{$T(\by) = T(\by_{obs})$,}
the first full conditional, $\pi(\bth|\by, T(\by)=T(\by_{obs}))$, is
\red{the same as} the full data posterior
$\pi(\bth|\by)$. 
\green{In this case, the condition $T(\by) = T(\by_{obs})$ is redundant.  This allows us to make use of conventional MCMC steps for this generation.  For} 
\sout{In the case of} typical regression models, algorithms abound.  Details of the \green{recommended} algorithms depend on details of
the prior distribution and sampling density.  As examples, a normal
prior distribution and normal likelihood in the regression setting
allow one to alternate conditional generations of $\sigma^2$ and $\bbeta$, and blocking the generation of $\bbeta$ generally
leads to quicker convergence and mixing \citep{liu1994}; a thick-tailed scale mixture of normal distributions
in the style of the 
hyper-$g/n$ prior \citep{liang2008} necessitates an additional stage for the sampler where the scale 
$g/n$ is generated; a thick-tailed sampling density such as a $t$ distribution can be handled with the addition of 
a scale parameter for each case and an extra stage where these scale parameters are generated.

%The other stage in the Gibbs sampler is the generation of complete data from the full conditional  $f(\by|\bth, T(\by)=T(\by_{obs}))$. %Condition \ref{supReal} above facilitates the extension of convergence proofs for the complete data algorithm to those for the incomplete data algorithm.  
%In this subsection, we focus exclusively this stage. 
%we begin with any conventional complete data algorithm.  In the case of typical regression models, these algorithms abound.  Details of the algorithm depend on details of
%the prior distribution and sampling density.  As examples, a normal
%prior distribution and normal likelihood in the regression setting
%allow one to alternate conditional generations of $\sigma^2$ and $\bbeta$, and blocking the generation of $\bbeta$ generally
%leads to quicker convergence and mixing \citep{liu1994}; a thick-tailed scale mixture of normal distributions
%in the style of the 
%hyper-$g/n$ prior \citep{liang2008} necessitates an additional stage for the sampler where the scale 
%$g/n$ is generated; a thick-tailed sampling density such as a $t$ distribution can be handled with the addition of 
%a scale parameter for each case and an extra stage where these scale parameters are generated.  
%The additional stage needed to implement the incomplete likelihood analysis via MCMC 
%is a generation of the complete data given statistic ($T(\by_{obs})$) and parameter ($\bth$).  Condition \ref{supReal} facilitates the extension of convergence proofs for the complete data algorithm to those for the incomplete data algorithm.  
%In this subsection, we focus exclusively on the additional stage
%where we generate $\by | \bth, T(\by)$.  
For a typical model and conditioning statistic, the second full conditional $f(\by|\bth, T(\by)=T(\by_{obs}))\blue{ =f(\by|\bth, \by\in\mc A)}$ is not available in closed form.  
As a consequence, we turn to Metropolis-Hastings \citep{hastings1970},
using the strategy of proposing full data $\by$ \green{\sout{in the sample
space of the distribution} for which $T(\by) = T(\by_{obs})$} and either accepting or rejecting the
proposal. Let $\by_p$ and  $\by_c$ represent the proposed and current
full data, respectively.  \green{In subsequent algorithms, we propose only from the set $\mathcal{A}$}. Denote the proposal distribution for $\by_{p}$ by $p(\by_p|\bth,T(\by_p) = T(\by_{obs})) \green{ = p(\by_p|\bth,\by_p \in \mathcal{A}) = p(\by_p | \bth)}$.  
\green{The last equality follows from the fact that our $p(\cdot | \bth)$ assigns probability one to the event $\{ \by_p \in \mathcal{A} \}$.  These equalities still hold if the dummy argument $\by_p$ is replaced with $\by_c$.  The conditional density is
\begin{eqnarray*}
f(\by | \bth, \by \in \mathcal{A}) & = & \frac{f(\by | \bth) I(\by \in \mathcal{A} | \by, \bth)}{\int_\mathcal{A} f(\by | \bth) d\by} \\
     & = & \frac{f(\by | \bth)}{\int_\mathcal{A} f(\by | \bth) d\by} 
\end{eqnarray*}
for $\by \in \mathcal{A}$.  This includes both $\by_p$ and $\by_c$.  
}

The Metropolis-Hastings acceptance probability  is the minimum of 1 and $R$ where,
\begin{eqnarray}
\label{MHRatio}
R & = & \frac{f(\by_p|\bth,\by_p \in \mathcal{A})}{f(\by_c|\bth,\by_c \in \mathcal{A})}  
                \green{\frac{p(\by_c|\bth, \by_c \in \mathcal{A})}{p(\by_p|\bth,\by_p \in \mathcal{A})}} \\
  & = & \frac{f(\by_p | \bth)}{\int_\mathcal{A} f(\by | \bth) d\by} \frac{\int_\mathcal{A} f(\by | \bth) d\by}{f(\by_c | \bth)} \frac{p(\by_c | \bth)}{p(\by_p | \bth)} \\
 & = & \frac{f(\by_p|\bth)}{f(\by_c|\bth)} \frac{p(\by_c|\bth)}{p(\by_p|\bth)} .  
\end{eqnarray}
\green{Note:  Above expression changed.  Check argument again.}
%In the above expression, the second line follows from the first for
%the following reason. The proposed \green{vector $\by$} is sampled from
%  $\mathcal{A}$ so that
%$T(\by_{p})=T(\by_{c})=T(\by_{obs})$. This implies 
%\green{\sout{$f(\by |\bth,
%T(\by)=T(\by_{obs}))=f(\by |\bth)\pi(\bth)/\red{f[?]}\pi(\bth, T(\by)=T(\by_{obs}))$}
%\begin{eqnarray}
%f(\by |\bth,\by \in \mathcal{A})) & = & \frac{f(\by | \bth) \pi(\bth) I(\by \in \mathcal{A})}{\int f(\by | \bth) \pi(\bth) I(\by \in \mathcal{A}) d\bth}
%\end{eqnarray}}
%where $\by$ represents either $\by_{p}$ or $\by_{c}$. %Further, the expressions $\pi(\bth)$ and $\red{f[?]}\pi(\bth, T(\by)=T(\by_{obs}))$ 
%\green{The expressions $\pi(\bth)$ and $\int f(\by | \bth) \pi(\bth) I(\by \in \mathcal{A})  d\bth$} are the same for $\by_{p}$ and $\by_{c}$ and cancel in the ratio. %Similar expansions can be made for the proposal distribution. 
%\green{Further, the indicator functions can be dropped since $\by_p$ and $\by_c$ will both fall in $\mathcal{A}$.}  
%\green{Note:  Check this argument carefully}

For the models we consider, evaluation of $f(\by | \bth)$ is straightforward.  Therefore, the difficulty in implementing this Metropolis-Hastings step manifests  itself in the ability to simulate proposals that guarantee $T(\by) = T(\by_{obs})$ and on the evaluation of the proposal density derived from the method used to simulate $\by$. We now discuss such an implementation method for the linear model in \eqref{LinearModel}.

\subsubsection{Construction of the proposal}
%\textcolor{blue}{some major revisions as indicated in the responseToReferees}
Our \green{computational} strategy is easiest to envision in a location-scale setting where 
the design matrix in model (\ref{LinearModel}) consists of a single column of ones.  Robust estimation techniques along
with model (\ref{LinearModel}) suggest a conditioning statistic $T(\by)$ which consists of
estimates of the scalars $\beta$ and $\sigma$, say $T(\by)=(b(X,\by),
s(X,\by))$.  It is not a simple matter to directly sample the required
proposal $\by$ satisfying $T(\by)=T(\by_{obs})$. However, 
  with equivariance/invariance properties of $T(\by)$, we will see in Theorem \ref{Transformation} below
    that it is possible to scale and shift any \sout{$\bz^{*}\in \mathbb R^{n}$} \green{$\bz^{*}$ which generates 
a positive scale estimate} to such a
    $\by$. Hence, to obtain $\by$ from $\mathcal A$, we proceed
  in two steps. First, a vector $\bz^*$ is sampled from a known
  distribution (examples of which are described later). This vector has summary statistic   
$T(\bz^*) = (b(X,\bz^*), s(X,\bz^*))$ which may not match the observed summary
$T(\by_{obs})$.  Second, $\bz^*$ is mapped into $\by =h(\bz^*)$
using Theorem \ref{Transformation} so that $T(\by)=T(\by_{obs})$. To
evaluate the proposal density $p(\by|\bth)$, we need to adjust
the known density of $\bz^*$, $p(\bz^*)$, with the Jacobian of
the transformation $h(\cdot)$ to $\by$. We choose the
distribution of the \green{\sout{intermediate}} data vector $\bz^*$ to have support
\green{on a subset of $\mathbb R^n$ so that}
the transformation to $\by$ is one-to-one and the Jacobian,
$\displaystyle \left|\frac{\partial h^{-1}(\by)}{\partial \by}\right|$, can be computed from first principles. It is the calculation of this Jacobian which is described in the section. We will repeatedly return to an artificial low-dimensional example in the location-scale setting to help visualize this calculation.

% To 
%obtain data $\by$ for which $T(\by) = T(\by_{obs})$, we proceed in two
%steps.  
%First, a vector $\bz^*$ is generated from a simple manifold with known
%density, for example, a uniform distribution
%on the surface of the unit sphere in ${\mathbb R}^{n-1}$ (alternatively from
%the model with current values of $\beta$ and $\sigma$).  This leads to 
%$T(\bz^*) = (b(X,\bz^*), s(X,\bz^*))$.  The vector $\bz^*$ is mapped into a vector
%$\by$ by rescaling and shifting appropriately to match the observed conditioning statistic.  For typical estimators, the appropriate scaling is 
%$s(X,\by_{obs}) / s(X,\bz^*)$ with the shift trailing along as needed to match $b(X,\by_{obs})$.  
%To evaluate the proposal density of $\by$,
%we need to adjust the density of $\bz^*$ with a Jacobian.  In the sequel, we use this artificially low-dimensional
%example to illustrate the method.

The strategy described in the previous paragraph extends to full-blown regression models.  Robust regression methods lead naturally to a conditioning statistic in the form of a classical M-estimator for $\bbeta$
and a companion estimator for $\sigma$.  We denote the resulting estimator which involves the covariates through
the design matrix and the response 
as $T(\by) = (\bb(X,\by), s(X,\by))$, with $\bb(X,\by) = (b_1(X,\by), \cdots,
b_p(X,\by))^\top$.  Simultaneous M-estimators have a number of standard properties \ref{asb}-\ref{scaleEq2Reg} which
prove useful in the sequel \citep{huber2009, maronna2006}.  
\begin{itemize}
\labitem{C3}{asb}$\bb(X,\by)$ is almost surely continuous and differentiable with respect to $\by$.  
\labitem{C4}{as} $s(X,\by)$ is almost surely positive, continuous, and differentiable with respect to $\by$.  
\labitem{C5}{regEq} $\bb(X,\by+X\bv)=\bb(X,\by)+\bv \ \ \text{for  all}\ \bv\in\mathbb{R}^p$. 
\labitem{C6}{scaleEqReg} $\bb(X,a\by)=a\bb(X,\by)\ \ \ \text{for all constants } a$.  
\labitem{C7}{regIn} $s(X,\by+X\bv)=s(X,\by) \ \ \text{for all}\ \bv\in\mathbb{R}^p$.  
\labitem{C8}{scaleEq2Reg} $s(X, a\by)=|a|s(X,\by) \ \ \text{for all constants } a$.  
\end{itemize}
Properties \ref{regEq} and \ref{scaleEqReg} of $\bb$ are called
\textit{regression} and \textit{scale equivariance},
respectively.  Properties \ref{regIn} and \ref{scaleEq2Reg} of $s$ are called \textit{regression invariance}
and \textit{scale equivariance}. 
Many other estimators satisfy these properties, and our subsequent results apply equally
well to them.  With more cumbersome statements, the upcoming results can be adjusted to handle 
a relaxation of \ref{as} that 
$s(X,\by_{obs}) > 0$ and $P(s(X,\by)>0)>0$.

>>>>

For the restricted likelihood model,  conditioning statistics are assumed to be of the form $T(\by) = (\bb(X, \by), s(X, \by))$ where $\bb(X, \by)= (b_1(X,\by), \dots,b_p(X,\by))^\top\in \mathbb R^{p}$ is an estimator for the regression coefficients and $s(X, \by)\in \{0\} \cup {\mathbb R}^+$ is an estimator of the scale. Throughout, observed data and summary statistic is denoted by $\by_{obs}$ and $T(\by_{obs})=(\bb(X, \by_{obs}), s(X, \by_{obs}))$, respectively. 
% In particular, we demonstrate the method using simultaneous M-estimators defined by \eqref{Mest} for the one-sample setting and easily extended to the linear model in \eqref{LinearModel}. The estimator of the \green{regression} coefficients is denoted by $\bb(X, \by)\in \mathbb R^{p}$ and that of the scale by \green{\sout{$s(X, \by)\in \mathbb R$} $s(X, \by)\in \{0\} \cup {\mathbb R}^+$} . Thus, $T(\by)=(\bb(X, \by), s(X, \by))$. 
Several conditions are imposed on the model and statistic to ensure validity of the MCMC algorithm:
\begin{itemize}
\labitem{C1}{fullRank} The $n \times p$ design matrix, $X$, whose $i^{th}$ row is $x_i^\top$, 
is of full column rank.  
\labitem{C2}{supReal} The $\epsilon_i$ are a random sample from some distribution which has a density with 
respect to Lebesgue measure on the real line and for which the support is the real line.  
\labitem{C3}{asb}$\bb(X,\by)$ is almost surely continuous and differentiable with respect to $\by$.  
\labitem{C4}{as} $s(X,\by)$ is almost surely positive, continuous, and differentiable with respect to $\by$.  
\labitem{C5}{regEq} $\bb(X,\by+X\bv)=\bb(X,\by)+\bv \ \ \text{for  all}\ \bv\in\mathbb{R}^p$. 
\labitem{C6}{scaleEqReg} $\bb(X,a\by)=a\bb(X,\by)\ \ \ \text{for all constants } a$.  
\labitem{C7}{regIn} $s(X,\by+X\bv)=s(X,\by) \ \ \text{for all}\ \bv\in\mathbb{R}^p$.  
\labitem{C8}{scaleEq2Reg} $s(X, a\by)=|a|s(X,\by) \ \ \text{for all constants } a$.  
\end{itemize}
Properties \ref{regEq} and \ref{scaleEqReg} of $\bb$ are called
\textit{regression} and \textit{scale equivariance},
respectively.  Properties \ref{regIn} and \ref{scaleEq2Reg} of $s$ are called \textit{regression invariance}
and \textit{scale equivariance}. 
Many estimators satisfy the above properties, including simultaneous M-estimators \citep{huber2009, maronna2006} for which the R package \texttt{brlm} (\texttt{github.com/jrlewi/brlm}) is available to implement the MCMC described here. Further software development is required to extend the MCMC implementation beyond these M-estimators. The package also implements the direct computational methods described in \cite{lewis2014}. These methods are effective in lower dimensional problems and were used in several of the examples in Section \ref{illustrations}.

@@

%The general style of algorithm that we present relies on the decomposition of the sampling density in (\ref{FullLikelihood})
%into one piece involving only $T(\by)$ and a second piece for the
%complete data $\by$ given $T(\by)$. 
%Relying on the modularity of MCMC algorithms, we construct a data augmented Gibbs sampler targeting $\pi(\bth, \by | T(\by)=T(\by_{obs}))$, the joint posterior of $\bth$ and $\by$ conditioned on the summary statistic $T(\by_{obs})$ for observed data vector $\by_{obs}$. The Gibbs sampler iteratively samples from the full conditionals written as $\pi(\bth|\by T(\by)=T(\by_{obs})$ and $f(\by|\bth, T(\by)=T(\by_{obs})$
% \textcolor{blue}{some major changes from here to the end of the section}
The general style of algorithm we present is a data augmented
MCMC targeting $f(\bth, \by |
T(\by)=T(\by_{obs}))$, the joint distribution of $\bth$ and the full
data given the summary statistic $T(\by_{obs})$. 
 The Gibbs sampler \citep{gelfand1990} iteratively samples from the
 full conditionals 
 \begin{enumerate}
 \item $\pi(\bth|\by, T(\by)=T(\by_{obs}))$ 
 \item $f(\by|\bth, T(\by)=T(\by_{obs}))$.  
 \end{enumerate}
When $\by$ has the summary statistic $T(\by) = T(\by_{obs})$,
the first full conditional is the same as the full data posterior $\pi(\bth|\by)$. In this case, the condition $T(\by) = T(\by_{obs})$ is redundant.  This allows us to make use of conventional MCMC steps for this generation.  For typical regression models, algorithms abound. Details of the recommended algorithms depend on details of
the prior distribution and sampling density and we assume this can be done \citep[see e.g.,][]{liu1994, liang2008}.

%As examples, a normal
%prior distribution and normal likelihood in the regression setting
%allow one to alternate conditional generations of $\sigma^2$ and $\bbeta$, and blocking the generation of $\bbeta$ generally
%leads to quicker convergence and mixing \citep{liu1994}; a thick-tailed scale mixture of normal distributions
%in the style of the hyper-$g/n$ prior \citep{liang2008} necessitates an additional stage for the sampler where the scale 
%$g/n$ is generated; a thick-tailed sampling density such as a $t$ distribution can be handled with the addition of 
%a scale parameter for each case and an extra stage where these scale parameters are generated.

%\green{Alternative algorithms substitute other generations for the full conditionals, such as the Metropolis-Hastings propose and accept/reject framework.  Our algorithms are based on the decomposition of the MCMC algorithm into conventional full data sampling steps where $\bth$ is generated given $\by$, and steps that allow us to complete the data set where $\by$ is generated given $\bth$ and $T(\by)$.  The latter generation relies on the conditioning set $\mathcal{A}=\{\by\in \mathbb{R}^n|T(\by)=T(\by_{obs})\}$.  The set $\mathcal{A}$ may contain more than one element (i.e., it is not degenerate) when $T(\by) : \by \rightarrow T(\by)$ is not a $1-1$ function.}

%The other stage in the Gibbs sampler is the generation of complete data from the full conditional  $f(\by|\bth, T(\by)=T(\by_{obs}))$. %Condition \ref{supReal} above facilitates the extension of convergence proofs for the complete data algorithm to those for the incomplete data algorithm.  
%In this subsection, we focus exclusively this stage. 
%we begin with any conventional complete data algorithm.  In the case of typical regression models, these algorithms abound.  Details of the algorithm depend on details of
%the prior distribution and sampling density.  As examples, a normal
%prior distribution and normal likelihood in the regression setting
%allow one to alternate conditional generations of $\sigma^2$ and $\bbeta$, and blocking the generation of $\bbeta$ generally
%leads to quicker convergence and mixing \citep{liu1994}; a thick-tailed scale mixture of normal distributions
%in the style of the 
%hyper-$g/n$ prior \citep{liang2008} necessitates an additional stage for the sampler where the scale 
%$g/n$ is generated; a thick-tailed sampling density such as a $t$ distribution can be handled with the addition of 
%a scale parameter for each case and an extra stage where these scale parameters are generated.  
%The additional stage needed to implement the incomplete likelihood analysis via MCMC 
%is a generation of the complete data given statistic ($T(\by_{obs})$) and parameter ($\bth$).  Condition \ref{supReal} facilitates the extension of convergence proofs for the complete data algorithm to those for the incomplete data algorithm.  
%In this subsection, we focus exclusively on the additional stage
%where we generate $\by | \bth, T(\by)$.  
For a typical model and conditioning statistic, the second full conditional $f(\by|\bth, T(\by)=T(\by_{obs}))$ %\blue{ =f(\by|\bth, \by\in\mc A)}$ 
is not available in closed form.  We turn to Metropolis-Hastings \citep{hastings1970},
using the strategy of proposing full data $\by \in \mathcal{A}:=\{\by \in \mathbb{R}^n | T(\by)=T(\by_{obs})\}$ from a well defined distribution with support $\mathcal{A}$ and either accepting or rejecting the
proposal. Let $\by_p, \by_c \in \mathcal{A}$ represent the proposed and current
full data, respectively. Denote the proposal distribution for $\by_{p}$ by $p(\by_p|\bth,T(\by_p) = T(\by_{obs})) = p(\by_p|\bth,\by_p \in \mathcal{A}) = p(\by_p | \bth)$.  The last equality follows from the fact that our $p(\cdot | \bth)$ assigns probability one to the event $\{ \by_p \in \mathcal{A} \}$.  These equalities still hold if the dummy argument $\by_p$ is replaced with $\by_c$.  The conditional density is
\begin{eqnarray*}
f(\by | \bth, \by \in \mathcal{A}) & = & \frac{f(\by | \bth) I(\by \in \mathcal{A} | \by, \bth)}{\int_\mathcal{A} f(\by | \bth) d\by} \\
     & = & \frac{f(\by | \bth)}{\int_\mathcal{A} f(\by | \bth) d\by} 
\end{eqnarray*}
for $\by \in \mathcal{A}$.  This includes both $\by_p$ and $\by_c$.  The Metropolis-Hastings acceptance probability  is the minimum of 1 and $R$ where,
\begin{eqnarray}
\label{MHRatio}
R & = & \frac{f(\by_p|\bth,\by_p \in \mathcal{A})}{f(\by_c|\bth,\by_c \in \mathcal{A})}  
                \frac{p(\by_c|\bth, \by_c \in \mathcal{A})}{p(\by_p|\bth,\by_p \in \mathcal{A})} \\
  & = & \frac{f(\by_p | \bth)}{\int_\mathcal{A} f(\by | \bth) d\by} \frac{\int_\mathcal{A} f(\by | \bth) d\by}{f(\by_c | \bth)} \frac{p(\by_c | \bth)}{p(\by_p | \bth)} \\
 & = & \frac{f(\by_p|\bth)}{f(\by_c|\bth)} \frac{p(\by_c|\bth)}{p(\by_p|\bth)} .  
\end{eqnarray}
%In the above expression, the second line follows from the first for
%the following reason. The proposed \green{vector $\by$} is sampled from
%  $\mathcal{A}$ so that
%$T(\by_{p})=T(\by_{c})=T(\by_{obs})$. This implies 
%\green{\sout{$f(\by |\bth,
%T(\by)=T(\by_{obs}))=f(\by |\bth)\pi(\bth)/\red{f[?]}\pi(\bth, T(\by)=T(\by_{obs}))$}
%f(\by |\bth,\by \in \mathcal{A})) & = & \frac{f(\by | \bth) \pi(\bth) I(\by \in \mathcal{A})}{\int f(\by | \bth) \pi(\bth) I(\by \in \mathcal{A}) d\bth}
%\end{eqnarray}}
%where $\by$ represents either $\by_{p}$ or $\by_{c}$. %Further, the expressions $\pi(\bth)$ and $\red{f[?]}\pi(\bth, T(\by)=T(\by_{obs}))$ 
%\green{The expressions $\pi(\bth)$ and $\int f(\by | \bth) \pi(\bth) I(\by \in \mathcal{A})  d\bth$} are the same for $\by_{p}$ and $\by_{c}$ and cancel in the ratio. %Similar expansions can be made for the proposal distribution. 
%\green{Further, the indicator functions can be dropped since $\by_p$ and $\by_c$ will both fall in $\mathcal{A}$.}  
%\green{Note:  Check this argument carefully}
For the models we consider, evaluation of $f(\by | \bth)$ is straightforward.  Therefore, the difficulty in implementing this Metropolis-Hastings step manifests  itself in the ability to both simulate from and evaluate $p(\by_p | \bth)$; the well defined distribution with support $\mathcal{A}$. We now discuss such an implementation method for the linear model in \eqref{LinearModel}.

\subsubsection{Construction of the proposal}
%\textcolor{blue}{some major revisions as indicated in the responseToReferees}
Our computational strategy relies on proposing $\by$ such that $T(\by) = T(\by_{obs})$ where $T(\cdot) = (\bb(X, \cdot), s(X, \cdot))$ satisfies the conditions \ref{asb}-\ref{scaleEq2Reg}. It is not a simple matter to do this directly, but with the specified conditions, it is possible to scale and shift any $\bz^{*}$ which generates 
a positive scale estimate to such a $\by$ via the following Theorem whose proof is in the appendix. 
\begin{theorem}
\label{Transformation}
Assume that conditions \ref{as}-\ref{scaleEq2Reg} hold.  Then, any vector $\bz^* \in \mathbb{R}^n$ with conditioning statistic
$T(\bz^*)$ for which $s(X,\bz^*) > 0$ can be transformed into $\by$ with conditioning statistic $T(\by) = T(\by_{obs})$ 
through the transformation 
\[
\by = h(\bz^*) := \frac{s(X,\by_{obs})}{s(X,\bz^*)} \bz^* + X\left(\bb(X,\by_{obs}) - \bb(X,\frac{s(X,\by_{obs})}{s(X,\bz^*)} \bz^*)\right) .  
\]
\end{theorem}
% (details are given in Theorem \ref{Transformation} below). Hence, to obtain $\by \in\mathcal A$, we proceed in two steps:
%\begin{enumerate}
%\item Sample $\bz^*\in \mathbb{R}^{n}$ from a known distribution (examples of which are described later).
%\item Transform $\bz^*$ to $\by =h(\bz^*)$ using Theorem \ref{Transformation} so that $T(\by)=T(\by_{obs})$. 
%\end{enumerate}
% To 
%obtain data $\by$ for which $T(\by) = T(\by_{obs})$, we proceed in two
%steps.  
%First, a vector $\bz^*$ is generated from a simple manifold with known
%density, for example, a uniform distribution
%on the surface of the unit sphere in ${\mathbb R}^{n-1}$ (alternatively from
%the model with current values of $\beta$ and $\sigma$).  This leads to 
%$T(\bz^*) = (b(X,\bz^*), s(X,\bz^*))$.  The vector $\bz^*$ is mapped into a vector
%$\by$ by rescaling and shifting appropriately to match the observed conditioning statistic.  For typical estimators, the appropriate scaling is 
%$s(X,\by_{obs}) / s(X,\bz^*)$ with the shift trailing along as needed to match $b(X,\by_{obs})$.  
%To evaluate the proposal density of $\by$,
%we need to adjust the density of $\bz^*$ with a Jacobian.  In the sequel, we use this artificially low-dimensional
%example to illustrate the method.  
%The strategy described in the previous paragraph extends to full-blown regression models.  Robust regression methods lead naturally to a conditioning statistic in the form of a classical M-estimator for $\bbeta$
%and a companion estimator for $\sigma$.  We denote the resulting estimator which involves the covariates through
%the design matrix and the response 
%as $T(\by) = (\bb(X,\by), s(X,\by))$, with $\bb(X,\by) = (b_1(X,\by), \cdots,
%b_p(X,\by))^\top$.  Simultaneous M-estimators have a number of standard properties \ref{asb}-\ref{scaleEq2Reg} which
%prove useful in the sequel \citep{huber2009, maronna2006}.  
%\begin{itemize}
%\labitem{C3}{asb}$\bb(X,\by)$ is almost surely continuous and differentiable with respect to $\by$.  
%\labitem{C4}{as} $s(X,\by)$ is almost surely positive, continuous, and differentiable with respect to $\by$.  
%\labitem{C5}{regEq} $\bb(X,\by+X\bv)=\bb(X,\by)+\bv \ \ \text{for  all}\ \bv\in\mathbb{R}^p$. 
%\labitem{C6}{scaleEqReg} $\bb(X,a\by)=a\bb(X,\by)\ \ \ \text{for all constants } a$.  
%\labitem{C7}{regIn} $s(X,\by+X\bv)=s(X,\by) \ \ \text{for all}\ \bv\in\mathbb{R}^p$.  
%\labitem{C8}{scaleEq2Reg} $s(X, a\by)=|a|s(X,\by) \ \ \text{for all constants } a$.  
%\end{itemize}
%Properties \ref{regEq} and \ref{scaleEqReg} of $\bb$ are called
%\textit{regression} and \textit{scale equivariance},
%respectively.  Properties \ref{regIn} and \ref{scaleEq2Reg} of $s$ are called \textit{regression invariance}
%and \textit{scale equivariance}. 
%Many other estimators satisfy these properties, and our subsequent results apply equally
%well to them.  With more cumbersome statements, the upcoming results can be adjusted to handle 
%a relaxation of \ref{as} that 
%%$s(X,\by_{obs}) > 0$ and $P(s(X,\by)>0)>0$.  
%To understand the Jacobian calculation, we start with the following Theorem whose proof is in the appendix. 
%%We have stated that any vector $\bz^{*} \in \mathbb{R}^n$ can be transformed to a proposal
%%$\by$ satisfying $T(\by) = T(\by_{obs})$.  The properties above ensure
%%that this can be done through the mechanism presented in the following theorem.  The proof of this and other results appear in the appendix.
%\begin{theorem}
%\label{Transformation}
%Assume that conditions \ref{as}-\ref{scaleEq2Reg} hold.  Then, any vector $\bz^* \in \mathbb{R}^n$ with conditioning statistic
%$T(\bz^*)$ for which $s(X,\bz^*) > 0$ can be transformed into $\by$ with conditioning statistic $T(\by) = T(\by_{obs})$ 
%through the transformation 
%\[
%\by = h(\bz^*) := \frac{s(X,\by_{obs})}{s(X,\bz^*)} \bz^* + X\left(\bb(X,\by_{obs}) - \bb(X,\frac{s(X,\by_{obs})}{s(X,\bz^*)} \bz^*)\right) .  
%\]
%\end{theorem}
Using the theorem, the general idea is to first start with a sample $\bz^*$ from a known distribution, say $p(\bz^*)$, and transform via $h(\cdot)$ to $\by \in \mathcal{A}$. The proposal density $p(\by|\bth)$ is then a change-of-variables adjustment on $p(\bz^*)$ derived from $h(\cdot)$.
%To evaluate the proposal density $p(\by|\bth)$, we need to adjust the known density of $\bz^*$, $p(\bz^*)$, with the Jacobian of the transformation $h(\cdot)$ to $\by$. We choose the
%distribution of $\bz^*$ to have support on a subset of $\mathbb R^n$ so that
%the transformation to $\by$ is one-to-one and the Jacobian, $\displaystyle \left|\frac{\partial h^{-1}(\by)}{\partial \by}\right|$, can be computed directly.  
In general however, the mapping $h()$ is many-to-one: for any $\bv\in \mathbb{R}^{n}$ and any $c\in \mathbb{R}^{+}$, $c\bz^{*} + X\bv$ map to the same $\by$. This makes the change-of-variables adjustment difficult.
%The set $\mathcal{A}$ typically does not lie in a linear space of dimension $n - p - 1$, and 
%so we must account for both the many-to-one nature of the mapping $h(\dot)$ and
%a Jacobian when deriving the proposal density.  
